\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-12-19 archive.tex
color-coded and cleaned up arrows

\begin{figure}
  \begin{tikzpicture}
    \renewcommand{\baselinestretch}{0.5}

%    \draw (0,0) rectangle (10,5);
    \node(t1)[align=left] at (1,-1) {
      \footnotesize +106 769-43$\to $106.769.43\\
      \footnotesize +83 973-831$\to $83.973.831\\
      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(}\subnode{f01}{${\color{orange}f_0}$}\mcode{  \codechar{.} \codechar{-}}\\
      \footnotesize \hspace{1.25cm}\mcode{(${\color{orange}f_0}$ \codechar{.} \codechar{ }}\\
      \footnotesize \hspace{1.4cm}\mcode{(cdr s)))}
    };
    \node(t2)[align=left] at ([xshift=2cm]t1.east) {
      \footnotesize Dream Coder$\to$Dr. Dream\\
      \footnotesize Andy Cencici$\to$Dr. Andy\\
      \footnotesize Jan Kotas$\to$Dr. Jan\\
      \footnotesize Nancy Jones$\to$Dr. Nancy\\
      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(${\color{orange}f_3}$ (${\color{orange}f_4}$ \codechar{ } s))}
    };
    \node(t3)[align=left] at ([xshift=2cm]t2.east) {
      \footnotesize Dream (Coder)$\to$Coder\\
      \footnotesize 1 (2) 3 4$\to$2\\
      \footnotesize the (dog) ran$\to$ran\\
      \footnotesize +817(33)99$\to$33\\
      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(${\color{orange}f_6}$ \codechar{(} \codechar{)} s)}
    };

@@

    \node[align=left,rotate=90] at (-1.5,-6.3) {Domain Specific Language (DSL)};

@@

    \invention{substitution}{1,-4}{
      ${\color{orange}f_0}($\mcode{s}$,$\mcode{a}$,$\mcode{b}$) \,=\, $\\
       \mcode{(map ($\lambda$ (x) (if (= x a)\\ b x)) s)}
    }{substitute chars}
    \vinvention{cat}{substitution}{
      ${\color{orange}f_1}($\mcode{a}$,$\mcode{b}$) \,=\, $\\
      \mcode{(fold ($\lambda$ (x y)\\ (cons x y))\\ a b))}
    }{concat strings}
    \vinvention{looping}{cat}{
       ${\color{orange}f_2}($\mcode{c}$,$\mcode{p}$,$\mcode{s}$) \,=\, $\\
      \mcode{(unfold l ($\lambda$ (x) (= c (car x))) p ($\lambda$ (z) (cdr z)))}
    }{loop until char}

    \hinvention{constant}{substitution}{
      ${\color{orange}f_3}($\mcode{s}$) \,=\, $\\
      \mcode{(${\color{orange}f_1}$ s string)}
    }{prepend constant}
    \vinvention{drop}{constant}{
      ${\color{orange}f_4}($\mcode{c}$) \,=\, $\\
      \mcode{(${\color{orange}f_2}$ c ($\lambda$ }\\\mcode{(z) (car z)))}
    }{prefix}
    \vinvention{suffix}{drop}{
      ${\color{orange}f_5}($\mcode{s}$,$\mcode{c}$) \,=\, $\\
      \mcode{(fold ($\lambda$ (x y) }\\\mcode{(cdr (if (= x c)}\\\mcode{ s y))) s s)}
    }{suffix}

    \hinvention{delimited}{constant}{
      \mcode{${\color{orange}f_6}($x$,$y$,$s$)\,=\,$}\\
      \mcode{(${\color{orange}f_4}$ x (${\color{orange}f_5}$ y c))}
    }{delimited substring}
    \draw[->] (delimited.west) -- (drop.east);
    \draw[->] (delimited.west) -- (suffix.east);

    \vinvention{appendSingle}{delimited}{
      \mcode{${\color{orange}f_7}($c$,$s$)\,=\,$}\\
      \mcode{(${\color{orange}f_1}$ (cons c nil) s)}}{append character}

\vinvention{surround}{appendSingle}{
      \mcode{${\color{orange}f_8}($s$)\,=\,$}\\
      \mcode{(cons \codechar{(} (${\color{orange}f_7}$ \codechar{)} s))}}{surround in parens}
    \draw[->] (surround.west) to[in=200,out=160] ([yshift=-0.25cm]appendSingle.west);
    \draw[->] (appendSingle.west) to[in=40,out=120] ([yshift=0.25cm]cat.east);

@@

    \begin{scope}[shift={(0,-15)}]
      \invention{filter}{0,0}{
        ${\color{orange}f_0} = $\mcode{($\lambda$ (p l)\\ (fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil l))} 
        }{higher-order `filter'}

      \invention{three}{[yshift=-\verticalSpacing]filter.center}{\footnotesize ${\color{orange}f_1} = $\\\mcode{(+ 1 (+ 1 1))}}{The number `3'}
      \invention{prefix}{[yshift=-\verticalSpacing]three.center}{\footnotesize ${\color{orange}f_2} = $\mcode{($\lambda$ (n l) (map ($\lambda$ (i) (index i l)) (range (+ 1 n))))}}{First $n$ elements}
      \invention{single}{[xshift=\horizontalSpacing]filter.center}{\footnotesize ${\color{orange}f_3} = $\mcode{($\lambda$ (z l) (fold ($\lambda$ (x y) (cons x y)) z l))}}{Append element}
      \invention{for}{[yshift=-\verticalSpacing]single.center}{\footnotesize ${\color{orange}f_4} = $\mcode{(+ 1 ${\color{orange}f_1}$)}}{The number `4'}
      \invention{Max}{[yshift=-\verticalSpacing]for.center}{\footnotesize ${\color{orange}f_5} = $\mcode{($\lambda$ (l) %
          (car (${\color{orange}f_0}$ ($\lambda$ (x) (nil? (${\color{orange}f_0}$ l ($\lambda$ (z) (> z x))))) l)))}}{Max of list}
      \invention{largest}{[xshift=\horizontalSpacing]single.center}{\footnotesize ${\color{orange}f_6} = $\mcode{($\lambda$ (l n) %
          (${\color{orange}f_5}$ (${\color{orange}f_0}$ ($\lambda$ (x) (> n (length (${\color{orange}f_0}$ ($\lambda$ (y) (> x y)) l)))) l)))}}{$n^{\text{th}}$ largest element}
      \vinvention{count}{largest}{\footnotesize ${\color{orange}f_7} = $\mcode{($\lambda$ (x l) %
          (length (${\color{orange}f_0}$ ($\lambda$ (y) (= x y)) l)))}}{count}
      \vinvention{weird}{count}{\footnotesize ${\color{orange}f_8} = $\mcode{($\lambda$ (p l) %
          (map ($\lambda$ (x) (if (p x) (+ 1 x) 0))))}}
                 {Higher-order incrementer}

      \draw[->] (for.west) -- (three.east);
      \draw[->] ([yshift=-0.5cm]largest.west) -- (Max.east);
      \draw[->] ([yshift=-0.5cm]largest.west) to[out=240,in=340] ([yshift=-0.8cm]filter.east);
      \draw[->] (Max.west) -- (filter.east);
      \draw[->] ([yshift=0.1cm]count.west) to[out=90,in=300] ([yshift=-0.3cm]filter.east);

      \node(t1)[align=left] at (0,3) {
        \mcode{[1 4 5 8]$\to$[5 8]}\\
        \mcode{[9 8 3 8]$\to$[9 8 8]}\\
        \mcode{[3 4 5 6]$\to$[5 6]}\\
        \mcode{[6 5 4 3]$\to$[6 5]}\\
        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(${\color{orange}f_0}$ ($\lambda$ (x)}\\\hspace{0.5cm}\mcode{ (> x ${\color{orange}f_4}$)) z)}
      };
      \node(t2)[align=left] at ([xshift=3.75cm]t1.center) {
        \mcode{[9 2 4]$\to$[4 2 9]}\\
        \mcode{[4 3]$\to$[3 4]}\\
        \mcode{[3 4 5 6]$\to$[6 5 4 3]}\\
        \mcode{[]$\to$[]}\\
        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(fold ($\lambda$ (x a)}\\\hspace{0.25cm}\mcode{ (${\color{orange}f_3}$ x a)) nil z)}
      };
      \node(t3)[align=left] at ([xshift=3.75cm]t2.center) {
        \mcode{[9 2 3]$\to$[2 3 9]}\\
        \mcode{[2 1]$\to$[1 2]}\\
        \mcode{[8 7 9 2 5]$\to$[2 5 7 8 9]}\\
        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(map ($\lambda$ (i)}\\\hspace{0.5cm}\mcode{ (${\color{orange}f_6}$ z i)) }\\\hspace{0.5cm}\mcode{(+ 1 (range (length z)))}
      };

>>>>

\begin{figure}
  \begin{tikzpicture}
    \renewcommand{\baselinestretch}{0.5}
    \renewcommand{\inventionColor}{orange}

%    \draw (0,0) rectangle (10,5);
    \node(t1)[align=left] at (1,-1) {
      \footnotesize +106 769-43$\to $106.769.43\\
      \footnotesize +83 973-831$\to $83.973.831\\
      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(}\subnode{f01}{${\color{\inventionColor}f_0}$}\mcode{  \codechar{.} \codechar{-}}\\
      \footnotesize \hspace{1.25cm}\mcode{(${\color{\inventionColor}f_0}$ \codechar{.} \codechar{ }}\\
      \footnotesize \hspace{1.4cm}\mcode{(cdr s)))}
    };
    \node(t2)[align=left] at ([xshift=2cm]t1.east) {
      \footnotesize Dream Coder$\to$Dr. Dream\\
      \footnotesize Andy Cencici$\to$Dr. Andy\\
      \footnotesize Jan Kotas$\to$Dr. Jan\\
      \footnotesize Nancy Jones$\to$Dr. Nancy\\
      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(${\color{\inventionColor}f_3}$ (${\color{\inventionColor}f_4}$ \codechar{ } s))}
    };
    \node(t3)[align=left] at ([xshift=2cm]t2.east) {
      \footnotesize Dream (Coder)$\to$Coder\\
      \footnotesize 1 (2) 3 4$\to$2\\
      \footnotesize the (dog) ran$\to$ran\\
      \footnotesize +817(33)99$\to$33\\
      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(${\color{\inventionColor}f_6}$ \codechar{(} \codechar{)} s)}
    };

@@

    \node[align=left,rotate=90] at (-1.5,-6.3) {Domain Specific Language (DSL)\\\\System discovers subroutines in {\color{\inventionColor}orange}};

@@

    \invention{substitution}{1,-4}{
      ${\color{\inventionColor}f_0}($\mcode{s}$,$\mcode{a}$,$\mcode{b}$) \,=\, $\\
       \mcode{(map ($\lambda$ (x) (if (= x a)\\ b x)) s)}
    }{substitute chars}
    \vinvention{cat}{substitution}{
      ${\color{\inventionColor}f_1}($\mcode{a}$,$\mcode{b}$) \,=\, $\\
      \mcode{(fold ($\lambda$ (x y)\\ (cons x y))\\ a b))}
    }{concat strings}
    \vinvention{looping}{cat}{
       ${\color{\inventionColor}f_2}($\mcode{c}$,$\mcode{p}$,$\mcode{s}$) \,=\, $\\
      \mcode{(unfold l ($\lambda$ (x) (= c (car x))) p ($\lambda$ (z) (cdr z)))}
    }{loop until char}

    \hinvention{constant}{substitution}{
      ${\color{\inventionColor}f_3}($\mcode{s}$) \,=\, $\\
      \mcode{(${\color{\inventionColor}f_1}$ s string)}
    }{prepend constant}
    \vinvention{drop}{constant}{
      ${\color{\inventionColor}f_4}($\mcode{c}$) \,=\, $\\
      \mcode{(${\color{\inventionColor}f_2}$ c ($\lambda$ }\\\mcode{(z) (car z)))}
    }{prefix}
    \vinvention{suffix}{drop}{
      ${\color{\inventionColor}f_5}($\mcode{s}$,$\mcode{c}$) \,=\, $\\
      \mcode{(fold ($\lambda$ (x y) }\\\mcode{(cdr (if (= x c)}\\\mcode{ s y))) s s)}
    }{suffix}

    \hinvention{delimited}{constant}{
      \mcode{${\color{\inventionColor}f_6}($x$,$y$,$s$)\,=\,$}\\
      \mcode{(${\color{\inventionColor}f_4}$ x (${\color{\inventionColor}f_5}$ y c))}
    }{delimited substring}
    \draw[->] (delimited.west) -- (drop.east);
    \draw[->] (delimited.west) -- (suffix.east);

    \vinvention{appendSingle}{delimited}{
      \mcode{${\color{\inventionColor}f_7}($c$,$s$)\,=\,$}\\
      \mcode{(${\color{\inventionColor}f_1}$ (cons c nil) s)}}{append character}

\vinvention{surround}{appendSingle}{
      \mcode{${\color{\inventionColor}f_8}($s$)\,=\,$}\\
      \mcode{(cons \codechar{(} (${\color{\inventionColor}f_7}$ \codechar{)} s))}}{surround in parens}
    \draw[->] (surround.west) to[in=200,out=160] ([yshift=-0.25cm]appendSingle.west);
    \draw[->] (appendSingle.west) to[in=40,out=120] ([yshift=0.25cm]cat.east);

@@

    \begin{scope}[shift={(0,-15)}]
      \invention{filter}{0,0}{
        ${\color{\inventionColor}f_0} = $\mcode{($\lambda$ (p l)\\ (fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil l))} 
        }{higher-order `filter'}

      \invention{three}{[yshift=-\verticalSpacing]filter.center}{\footnotesize ${\color{\inventionColor}f_1} = $\\\mcode{(+ 1 (+ 1 1))}}{The number `3'}
      \invention{prefix}{[yshift=-\verticalSpacing]three.center}{\footnotesize ${\color{\inventionColor}f_2} = $\mcode{($\lambda$ (n l) (map ($\lambda$ (i) (index i l)) (range (+ 1 n))))}}{First $n$ elements}
      \invention{single}{[xshift=\horizontalSpacing]filter.center}{\footnotesize ${\color{\inventionColor}f_3} = $\mcode{($\lambda$ (z l) (fold ($\lambda$ (x y) (cons x y)) z l))}}{Append element}
      \invention{for}{[yshift=-\verticalSpacing]single.center}{\footnotesize ${\color{\inventionColor}f_4} = $\mcode{(+ 1 ${\color{\inventionColor}f_1}$)}}{The number `4'}
      \invention{Max}{[yshift=-\verticalSpacing]for.center}{\footnotesize ${\color{\inventionColor}f_5} = $\mcode{($\lambda$ (l) %
          (car (${\color{\inventionColor}f_0}$ ($\lambda$ (x) (nil? (${\color{\inventionColor}f_0}$ l ($\lambda$ (z) (> z x))))) l)))}}{Max of list}
      \invention{largest}{[xshift=\horizontalSpacing]single.center}{\footnotesize ${\color{\inventionColor}f_6} = $\mcode{($\lambda$ (l n) %
          (${\color{\inventionColor}f_5}$ (${\color{\inventionColor}f_0}$ ($\lambda$ (x) (> n (length (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (> x y)) l)))) l)))}}{$n^{\text{th}}$ largest element}
      \vinvention{count}{largest}{\footnotesize ${\color{\inventionColor}f_7} = $\mcode{($\lambda$ (x l) %
          (length (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (= x y)) l)))}}{count}
      \vinvention{weird}{count}{\footnotesize ${\color{\inventionColor}f_8} = $\mcode{($\lambda$ (p l) %
          (map ($\lambda$ (x) (if (p x) (+ 1 x) 0))))}}
                 {Higher-order incrementer}

      \draw[->] (for.west) -- (three.east);
      \draw[->] ([yshift=-0.0cm]largest.west) -- (Max.east);
      \draw[->] ([yshift=-0.cm]largest.west) to[out=240,in=330] ([yshift=-0.1cm]filter.east);
      \draw[->] (Max.west) -- (filter.east);
      \draw[->] ([yshift=0.1cm]count.west) -- ([yshift=-0.3cm]filter.east);

      \node(t1)[align=left] at (0,3) {
        \mcode{[1 4 5 8]$\to$[5 8]}\\
        \mcode{[9 8 3 8]$\to$[9 8 8]}\\
        \mcode{[3 4 5 6]$\to$[5 6]}\\
        \mcode{[6 5 4 3]$\to$[6 5]}\\
        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(${\color{\inventionColor}f_0}$ ($\lambda$ (x)}\\\hspace{0.5cm}\mcode{ (> x ${\color{\inventionColor}f_4}$)) z)}
      };
      \node(t2)[align=left] at ([xshift=3.75cm]t1.center) {
        \mcode{[9 2 4]$\to$[4 2 9]}\\
        \mcode{[4 3]$\to$[3 4]}\\
        \mcode{[3 4 5 6]$\to$[6 5 4 3]}\\
        \mcode{[]$\to$[]}\\
        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(fold ($\lambda$ (x a)}\\\hspace{0.25cm}\mcode{ (${\color{\inventionColor}f_3}$ x a)) nil z)}
      };
      \node(t3)[align=left] at ([xshift=3.75cm]t2.center) {
        \mcode{[9 2 3]$\to$[2 3 9]}\\
        \mcode{[2 1]$\to$[1 2]}\\
        \mcode{[8 7 9 2 5]$\to$[2 5 7 8 9]}\\
        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(map ($\lambda$ (i)}\\\hspace{0.5cm}\mcode{ (${\color{\inventionColor}f_6}$ z i)) }\\\hspace{0.5cm}\mcode{(+ 1 (range (length z)))}
      };
